Cognitive Principles and Individual Differences \\ in Human Syllogistic Reasoning

### 2019-11-18 paper/paper-v3.tex
made the (hopfully very final changes

Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
various experimental studies have shown that human reasoning systematically deviates 
from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}).
% Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
% However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
% On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
% in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
% % as they were mostly theoretical and did not apply to real case studies.
% However, only recently an attempt was made 
% this expertise was not consulted, i.e. there was no communication among the relevant fields. 
Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}. 
In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~the suggestion by Ragni and Stolzenburg~\cite{ragni:stolzenburg:2015} as a starting point:
We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.

% Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions. 
% Only then, the cognitve adequacy of that theory can be compared to others.
% 
% Formal methods can offer the required tools to implement such cognitive theories. This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy. 
% A major drawback of 
The results of the meta-study on syllogistic reasoning by Khemlani and Johnson-Laird~\cite{khemlani:2012} show that the answers given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
% Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities. 
% As recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
% This implies that \textit{the} human reasoner does not exist, instead there 
% seem to exist various (human) reasoning clusters.

@@

each of which is a statement about properties over terms which can have one of the classical quantifiess in Table~\ref{table:moods} (called \emph{moods}). The pair of syllogistic premises can have four different orders as shown in Table~\ref{t:2} (called \emph{figures}). Based on these moods and figures, 64 different pairs of syllogistic premises can be constructed.

@@

    \begin{minipage}{0.6\textwidth}

@@

\begin{array}{@{\hspace{0.0cm}}l@{\hspace{0.1cm}}G@{\hspace{0.1cm}}l@{\hspace{0.3cm}}c}\toprule

@@

	\caption{The four moods and their

@@

    \begin{minipage}{0.45\textwidth}

@@

	\caption{The four figures.\label{t:2}}

@@

was concluded by only 16\,\%, which is exactly the threshold that the answer had not been chosen randomly~\cite{khemlani:2012}.

@@

% i.e.\ the data summed up with respect to the individual answer possibilities.

@@

% these answers excluded each other. In particular, it can be assumed that participants were at least minimally consistent in the sense that

@@

Usually in these experiments, participants are supposed to give only one answer, thus all three (significant) answers

@@

% The ones who answered \emph{No valid conclusion}
% were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{}

A major drawback of the suggested cognitive theories so far is that only the participants'
aggregated data was considered: These theories did not account for differences
among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities. In contrast, non-aggregated data records the individual participants' response patterns, i.e.\ the complete set of responses that are given by each participant.

As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditionals differently.
This implies that \textit{the} human reasoner does not exist, instead there 
seem to exist various (human) reasoning clusters. The identification of these differences and the specification of these clusters is central 
for the development of an adequate cognitive theory. Only recently approaches have been proposed that account for individual differences (or clusters) in syllogistic reasoning (cf.~\cite{khemlani:2016,declare:2017}).

%  The results show that some but not all of the proposed clusters are well reflected in the individual reasoning patterns.
% More interestingly are the following two observations:
% The dataset in~\cite{khemlani:2016} provides an additional interesting information: 
% Interestingly, when comparing the changes of results of each individual participant
% from the first to the second week, it seems that their answers in the second week fits better the classical logic derivations. 
% We argue that these reasoners can be characterized by a set of cognitive principles, motivated from studies made in Cognitive Science.
% First attempts to propose a model that distinguishes between human reasoners have been proposed in~\cite{khemlani:2016}. 
% The authors identify three different clusters of reasoners labeled as intuitive, intermediate and deliberative. They propose an approaches based on the MReasoner and suggest to   ...
% We evaluate the non-aggregated data provided by~\cite{khemlani:2016} and taking the clusters suggested in~\cite{declare:2017} as starting point. The non-aggregated data consists of two dataset for 20 participants, who carried out the same syllogistic reasoning task twice with a interval of one week.
%  The evaluation shows that some but not all of the proposed clusters in~\cite{declare:2017} are well reflected in the individual reasoning patterns.
% Interestingly, when comparing the changes of results of each individual participant
% from the first to the second week, it seems that their answers in the second week fits better the classical logic derivations. 
% We formulate the hypothesis that the individual clusters do not represent different human reasoners, but different stages of reasoning. Depending on the persistence and the
% previous training of the individual participant, ...

Here, we focus on a qualitative assessment of the clustering approach proposed in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.

\section{Clustering by Principles under WCS} \label{sect:wcsclustering}

Clustering by principles under WCS was first presented in~\cite{declare:2017};
the authors conjectured that reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches by Khemlani an Johnson-Laird~\cite{khemlani:2012}.
% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
Next, the respective models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
 The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.

@@

\node[anchor=west]  at (0,9.2) {\textbf{Cognitive Principles} presuppositions that humans make while reasoning};
\node[anchor=west]  at (0,8.7) {\textbf{Heuristic Strategies} reasoning based on structure in patterns or biases};

@@

\end{tikzpicture}
\caption{The system that participated at the Syllogism Challenge 2017\label{overview:system:syllchal}.}

@@

The suggested computational logic approach for the formal representation and reasoning of these clusters,
is the Weak Completion Semantics (WCS)~\cite{hk:2009a,hk:2009b}
which has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
% WCS is the non-monotonic computational logic approach to consider
% weakly completed logic programs, to compute their least models
% under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.

@@

We introduce an extension of WCS first presented in~\cite{dietz:hoelldobler:pereira:2017}. 
\textit{Contextual (logic) program clauses}
are expressions of the forms $A  \leftarrow  L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ (called
\textit{rules}), $A \leftarrow \top$ (called \textit{facts}), $A \leftarrow \bot$ (called
\textit{negative assumptions})\footnote{Under WCS, the negative
assumption will become $A \leftrightarrow \bot$, which, however, can be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
\textit{unknown assumptions}), where
$A$ is an atom and~$L_i$ with~$1 \leq i \leq m+p$, is a literal.
$A$ is called \textit{head}
and $L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ as well as~$\top, \bot$
and $\udf$,
standing for \textit{true}, \textit{false} and \textit{unknown} respectively, are
called \textit{body} of the corresponding clauses.  A
\textit{contextual program}, denoted by $\mathcal P$,
is a finite set of contextual program clauses.
We restrict terms to be either constants or variables only, hence we consider so-called \defname{data logic programs}.
If a clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
% For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
A \defname{ground clause} is a clause that does not contain variables.
$\Ground \CalP$ denotes the set of all ground clauses of $\CalP$.
An atom $A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a clause with head $A$.
The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a clause
occurring in } \Ground \CalP\}$.
$A$ is \textit{undefined in $\Ground \CalP$} iff $A$ is not defined in $\Ground \CalP$.
The set of all atoms that are undefined in~$\Ground \CalP$ is denoted by~$\ud(\CalP)$.
Consider the following transformation for a given $\CalP$:
1.\ For each atom~$A$ defined in $\Ground \CalP$, replace
  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
The obtained ground set of equivalences is called the \textit{weak completion}
of~$\CalP$ or~$\WC\CalP$.
% $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
% contains a negative assumption with head $A$, $\Ground \CalP$ does not
% contain an unknown assumption with head $A$, and $\Def(A,\CalP) = \emptyset$.
We omit the word \textit{contextual} when we refer to programs, if not stated otherwise.

@@

We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
extended with the $\cxt$ connective, for which the
corresponding truth values are $\top$, $\bot$ and $\udf$, meaning \textit{true}, \textit{false}
and \textit{unknown}, respectively.
% As it seems,  $\cxt$ corresponds to {\L}ukasiewicz necessity operator~\cite{}, where the difference is, that $\cxt$ is restricted to be applicable to literals only.
A \defname{three-valued interpretation}~$I$ is a mapping from the set of
logical formulas to the set
of truth values~$\{ \true, \bot, \udf\}$, represented as a pair~$I~=~\langle I^\true,I^\bot \rangle$ of two disjoint sets of atoms:  $I^\true =  \{A \mid A \mbox{ is mapped to } \true \mbox{ under } I\}$ and $I^\bot = \{A \mid A \mbox{ is mapped to } \bot \mbox{ under } I \}$.
Atoms which do not occur in~$I^\true \cup I^\bot$ are mapped to~$\udf$.
The truth value of a given formula
under $I$ is determined according to the truth tables in
Table~\ref{tab:3vldluka}.
\footnote{Thanks to the
 anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiwicz's necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
 }

@@

\begin{table}[t]
\[\small
\begin{array}[c]{@{\hspace{0mm}}c|c}
F & \neg F\\ \midrule
\top & \bot \\
\bot & \top \\
\udf & \udf
\end{array}
\;\
\begin{array}{c|lll}
 \wedge & \top & \udf & \bot \\
\midrule
\top & \top & \udf & \bot \\
\udf & \udf & \udf & \bot \\
\bot & \bot & \bot & \bot
 \end{array}
\;\
\begin{array}{c|lll}
 \vee & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \top & \udf & \udf \\
\bot & \top & \udf & \bot
 \end{array}
\;\
\begin{array}{c|lll}
\leftarrow & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \udf & \top & \top \\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|ccc}
  \leftrightarrow  & \top & \udf & \bot \\
  \midrule
\top & \top & \udf & \bot \\
\udf & \udf & \top & \udf\\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|c}
 L & \cxt(L)\\
\midrule
\top & \top \\
\bot & \bot  \\
\udf & \bot
\end{array}
\]
\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt$. $L$ is a literal and $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
and \textit{unknown}, respectively.\label{tab:3vldluka}}
\end{table}

@@

Consider the following semantic operator introduced by Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
\rangle$ be an interpretation. $\Phi_\CalP (I) =
\langle J^\top,J^\bot\rangle$, where
\[
\begin{array}{lcl}
J^\top & = & \{ A \mid A \leftarrow \Body \in \Def(A,\CalP)
\mbox{ and
} \Body \mbox{ is \True under } \langle I^\top,I^\bot \rangle \} \\
J^\bot & = & \{ A \mid \Def(A,\CalP) \neq \emptyset \mbox{ and }\\
&& \rule{7.5mm}{0mm}\Body \mbox{ is \False under } \langle
I^\top,I^\bot \rangle \mbox{ for all }  A \leftarrow \Body \in \Def(A,\CalP)\}.
\end{array}\Shrink
\]
The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
always has a least model under  {\L}ukasiewicz logic,
which can be computed by the least fixed point of~$\Svlp$~\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.

The weak completion of contextual
programs might have more than one minimal model, and it is not guaranteed that 
the least fixed point of $\Svlp$ always exists. However,
for programs without cycles, there always exists a unique supported model,
which corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}. This model is \textit{supported} in the sense that 
it will always be computed by $\Svlp$, independent from the interpretation we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$:
% As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, 
% we define 
$\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
For all programs considered in the sequel, $\CalM_\CalP$ exists.

@@

\section{Evaluation} \label{sect:aggregateddata}

\subsection{On aggregated Data}

>>>>

Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
various experimental studies have shown that human reasoning systematically deviates 
from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}).
% Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
% However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
% On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
% in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
% % as they were mostly theoretical and did not apply to real case studies.
% However, only recently an attempt was made 
% this expertise was not consulted, i.e. there was no communication among the relevant fields. 
Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}. 
In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~the suggestion by Ragni and Stolzenburg~\cite{ragni:stolzenburg:2015} as a starting point:
We aim for a theory in which (i) all processes are comprehensible, (ii) the underlying assumptions represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.

% Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions. 
% Only then, the cognitve adequacy of that theory can be compared to others.
% 
% Formal methods can offer the required tools to implement such cognitive theories. This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy. 
% A major drawback of 
The results of the meta-study on syllogistic reasoning by Khemlani and Johnson-Laird~\cite{khemlani:2012} show that the responses given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
% Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual response possibilities. 
% As recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
% This implies that \textit{the} human reasoner does not exist, instead there 
% seem to exist various (human) reasoning clusters.

@@

each of which is a statement about properties over terms which can have one of the classical quantifiers in Table~\ref{table:moods} (called \emph{moods}). The pair of syllogistic premises can have four different orders as shown in Table~\ref{t:2} (called \emph{figures}). Based on these moods and figures, 64 different pairs of syllogistic premises can be constructed.

@@

    \begin{minipage}{0.65\textwidth}

@@

\begin{array}{@{\hspace{0.0cm}}l@{\hspace{0.4cm}}G@{\hspace{0.1cm}}l@{\hspace{0.3cm}}c}\toprule

@@

	\caption{The classical moods and their

@@

    \begin{minipage}{0.34\textwidth}

@@

	\caption{The classical figures.\label{t:2}}

@@

was concluded by only 16\,\%, which is exactly the threshold that the response had not been chosen randomly~\cite{khemlani:2012}.

@@

% i.e.\ the data summed up with respect to the individual response possibilities.

@@

% these responses excluded each other. In particular, it can be assumed that participants were at least minimally consistent in the sense that

@@

Usually in these experiments, participants are supposed to give only one response, thus all three responses above

@@

% The ones who responseed \emph{No valid conclusion}
% were not the same ones who responseed \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{}

A major drawback of most cognitive theories is that only the participants'
aggregated data was considered: These theories did not account for differences
among individuals, i.e.\ they only considered the data summed up with respect to the individual response possibilities. In contrast, non-aggregated data records the individual participants' response patterns, i.e.\ the complete set of responses that are given by each participant.

As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditionals differently.
This implies that \textit{the} human reasoner does not exist, instead there 
seem to exist various groups of reasoners, i.e.\ reasoning clusters. The identification of these differences and the specification of these clusters is central 
for the development of an adequate cognitive theory. Only recently approaches have been proposed that account for individual differences (or clusters) in syllogistic reasoning (cf.~\cite{khemlani:2016,declare:2017}).

%  The results show that some but not all of the proposed clusters are well reflected in the individual reasoning patterns.
% More interestingly are the following two observations:
% The dataset in~\cite{khemlani:2016} provides an additional interesting information: 
% Interestingly, when comparing the changes of results of each individual participant
% from the first to the second week, it seems that their responses in the second week fits better the classical logic derivations. 
% We argue that these reasoners can be characterized by a set of cognitive principles, motivated from studies made in Cognitive Science.
% First attempts to propose a model that distinguishes between human reasoners have been proposed in~\cite{khemlani:2016}. 
% The authors identify three different clusters of reasoners labeled as intuitive, intermediate and deliberative. They propose an approaches based on the MReasoner and suggest to   ...
% We evaluate the non-aggregated data provided by~\cite{khemlani:2016} and taking the clusters suggested in~\cite{declare:2017} as starting point. The non-aggregated data consists of two dataset for 20 participants, who carried out the same syllogistic reasoning task twice with a interval of one week.
%  The evaluation shows that some but not all of the proposed clusters in~\cite{declare:2017} are well reflected in the individual reasoning patterns.
% Interestingly, when comparing the changes of results of each individual participant
% from the first to the second week, it seems that their responses in the second week fits better the classical logic derivations. 
% We formulate the hypothesis that the individual clusters do not represent different human reasoners, but different stages of reasoning. Depending on the persistence and the
% previous training of the individual participant, ...

Here, we provide a qualitative assessment of the \textit{clustering by principles} approach proposed in~\cite{declare:2017},
that performed well on aggregated data, and investigate whether these clusters are also adequate with respect to the non-aggregated dataset given by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach, the cognitive principles and their formalization. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.

\section{Clustering by Principles} \label{sect:wcsclustering}

Clustering by principles was first presented in~\cite{declare:2017}, which is an extension of
the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}.
The authors in~\cite{declare:2017} conjectured that reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when conclusions are not derived by reasoning but instead based on structural patterns or {biases}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches by Khemlani an Johnson-Laird~\cite{khemlani:2012}.
% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified through sets of underlying cognitive principles and heuristic strategies.
After that, these sets were translated into logic programs (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the reasoning clusters.
Next, the respective models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
After that, for each cluster, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
 The final response distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.

@@

\node[anchor=west]  at (0,9.2) {\textbf{Cognitive Principles} presuppositions that humans make while reasoning};
\node[anchor=west]  at (0,8.7) {\textbf{Heuristic Strategies} \hspace{0.05cm} reasoning based on structure in patterns or biases};

@@

\end{tikzpicture}
\caption{The methodology of \textit{clustering by principles}\label{overview:system:syllchal}.}

@@

The suggested computational logic approach for the formal representation and reasoning of these clusters,
is the Weak Completion Semantics (WCS)~\cite{hk:2009a,hk:2009b}
which has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}
and will be introduced in the following.
% WCS is the non-monotonic computational logic approach to consider
% weakly completed logic programs, to compute their least models
% under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.

@@

We introduce an extension of WCS first presented in~\cite{dietz:hoelldobler:pereira:2017}. 
\textit{Contextual (logic) program clauses}
are expressions of the forms $A  \leftarrow  L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ (called
\textit{rules}), $A \leftarrow \top$ (called \textit{facts}), $A \leftarrow \bot$ (called
\textit{negative assumptions})\footnote{Under WCS, the negative
assumption will become $A \leftrightarrow \bot$, which, however, can be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
\textit{unknown assumptions}), where
$A$ is an atom and~$L_i$ with~$1 \leq i \leq m+p$, is a literal.
$A$ is called \textit{head}
and $L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ as well as~$\top, \bot$
and $\udf$,
standing for \textit{true}, \textit{false} and \textit{unknown} respectively, are
called \textit{body} of the corresponding clauses.  A
\textit{contextual program}, denoted by $\mathcal P$,
is a finite set of contextual program clauses.
We restrict terms to be either constants or variables only, hence we consider so-called \defname{data logic programs}.
If a clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
% For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
A \defname{ground clause} is a clause that does not contain variables.
$\Ground \CalP$ denotes the set of all ground clauses of $\CalP$.
An atom $A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a clause with head $A$.
The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a clause
occurring in } \Ground \CalP\}$.
$A$ is \textit{undefined in $\Ground \CalP$} iff $A$ is not defined in $\Ground \CalP$.
The set of all atoms that are undefined in~$\Ground \CalP$ is denoted by~$\ud(\CalP)$.
Consider the following transformation for a given $\CalP$:
1.\ For each atom~$A$ defined in $\Ground \CalP$, replace
  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
The obtained ground set of equivalences is called the \textit{weak completion}
of~$\CalP$ or~$\WC\CalP$.
% $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
% contains a negative assumption with head $A$, $\Ground \CalP$ does not
% contain an unknown assumption with head $A$, and $\Def(A,\CalP) = \emptyset$.
In the sequel, we omit the word \textit{contextual} when we refer to programs, if not stated otherwise.

@@

We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
extended with the $\cxt$ connective, for which the
corresponding truth values are $\top$, $\bot$ and $\udf$, meaning \textit{true}, \textit{false}
and \textit{unknown}, respectively.
% As it seems,  $\cxt$ corresponds to {\L}ukasiewicz necessity operator~\cite{}, where the difference is, that $\cxt$ is restricted to be applicable to literals only.
A \defname{three-valued interpretation}~$I$ is a mapping from the set of
logical formulas to the set
of truth values~$\{ \true, \bot, \udf\}$, represented as a pair~$I~=~\langle I^\true,I^\bot \rangle$ of two disjoint sets of atoms:  $I^\true =  \{A \mid A \mbox{ is mapped to } \true \mbox{ under } I\}$ and $I^\bot = \{A \mid A \mbox{ is mapped to } \bot \mbox{ under } I \}$.
Atoms which do not occur in~$I^\true \cup I^\bot$ are mapped to~$\udf$.
The truth value of a given formula
under $I$ is determined according to the truth tables in
Table~\ref{tab:3vldluka}.

@@

\begin{table}[t]
\[\small
\begin{array}[c]{@{\hspace{0mm}}c|c}
F & \neg F\\ \midrule
\top & \bot \\
\bot & \top \\
\udf & \udf
\end{array}
\;\
\begin{array}{c|lll}
 \wedge & \top & \udf & \bot \\
\midrule
\top & \top & \udf & \bot \\
\udf & \udf & \udf & \bot \\
\bot & \bot & \bot & \bot
 \end{array}
\;\
\begin{array}{c|lll}
 \vee & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \top & \udf & \udf \\
\bot & \top & \udf & \bot
 \end{array}
\;\
\begin{array}{c|lll}
\leftarrow & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \udf & \top & \top \\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|ccc}
  \leftrightarrow  & \top & \udf & \bot \\
  \midrule
\top & \top & \udf & \bot \\
\udf & \udf & \top & \udf\\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|c}
 L & \cxt(L)\footnotemark\\
\midrule
\top & \top \\
\bot & \bot  \\
\udf & \bot
\end{array}
\]
\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt$. $L$ is a literal and $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
and \textit{unknown}, respectively.\label{tab:3vldluka}}
\end{table}
\footnotetext{Thanks to the
 anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiwicz's necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
 }

@@

Consider the following semantic operator introduced by Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
\rangle$ be an interpretation. $\Phi_\CalP (I) =
\langle J^\top,J^\bot\rangle$, where
\[
\begin{array}{lcl}
J^\top & = & \{ A \mid A \leftarrow \Body \in \Def(A,\CalP)
\mbox{ and
} \Body \mbox{ is \True under } \langle I^\top,I^\bot \rangle \} \\
J^\bot & = & \{ A \mid \Def(A,\CalP) \neq \emptyset \mbox{ and }\\
&& \rule{7.5mm}{0mm}\Body \mbox{ is \False under } \langle
I^\top,I^\bot \rangle \mbox{ for all }  A \leftarrow \Body \in \Def(A,\CalP)\}.
\end{array}\Shrink
\]
The weak completion of non-contextual programs\footnote{Non-contextual programs are programs that do not contain rules in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
always has a least model under  {\L}ukasiewicz logic,
which corresponds to the least fixed point of~$\Svlp$~\cite{hk:2009a}. 
The interested reader is referred to~\cite{dietz:hoelldobler:wernhard:2014} for the correspondence to other 
logic programming approaches.

Different than for non-contextual programs, the weak completion of contextual
programs might have more than one minimal model. Furthermore, if $\CalP$ is a contextual program, the least fixed point of $\Svlp$ might not even exist. However,
for contextual programs without cycles, there always exists a unique supported model,
which corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}. This model is \textit{supported} in the sense that 
it will always be computed by $\Svlp$, independent from the interpretation we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$.
As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, 
we define 
$\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
For all programs considered in the sequel, $\CalM_\CalP$ exists.

@@

\section{Evaluation}

\subsection{On aggregated Data}\label{sect:aggregateddata}
